What We Eval in the Shadows

### 2021-04-16 paper/main.tex
grammarly's proofreading for Section 3 (Methodoloy)

Our corpus is assembled from three sources: the \emph{Base} libraries that are
bundled with the language implementation, packages hosted on \emph{CRAN} and
scripts from \emph{Kaggle}.

\mypara{Base.} The dataset contains \BasePackages libraries performing basic
arithmetics, statistics, and operating system functionalities. These libraries
are bundled with R and executed pervasively. Together they contain
\BaseEvalCallSites call sites to \eval in \BaseFunsWithEvals functions. Some of these
functions provide basic functionalities such as package loading and thus there
is hardly any code which does not invoke one of these.

\mypara{CRAN.} The Comprehensive R Archive Network
({\small \url{cran.r-project.org}}) is the largest curated repository of R
packages. It hosts \CranAvailablePackagesRnd packages with 6 new ones submitted
daily~\cite{Ligges2017}. Packages are authored by experienced developers and
abide by well-formedness rules automatically checked on each commit.
Each package comes with sample data. There are three sources of
runnable code in a package: \emph{tests, examples} and \emph{vignettes} --
respectively, unit tests, code snippets from the documentation, and long-form
use-cases written in RMarkdown. Examples and vignettes can be turned into
scripts by extracting the relevant code. From \CranPackages packages we
extracted \CranRunnableScripts scripts that contain \CranRunnableCode lines of
code (\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd in
vignettes and \CranRunnableCodeTestsRnd in tests).

@@

The analysis presented in this has been automated by a pipeline that acquires
packages, extract scripts, executes them, traces their behavior and summarizes
observations. Figure~\ref{fig:pipeline} shows the main steps along with their
running time, data size, and number of elements manipulated. Timings are from a
cluster of three 2.3GHz Intel Xeon 6140 servers, with 72 cores and 256GB of RAM,
and shared OCFS network storage. The pipeline steps are:

@@

\medskip
\begin{compactenum}
\item \emph{Download.} Packages are downloaded from CRAN. For scripts, a web
  crawler retrieves code and the Kaggle command line tool gets data.
  Installation is complicated by native dependencies which are not properly
  documented and thus hard to automatically resolve.
\item \emph{Extract.} Given installed packages and scripts, the next step is to
  create runnable programs. The \genthat tool helpfully extracts all runnable
  code snippets from a package and turns each of these into a self-standing
  program~\cite{issta18}. Some Kaggle kernels are already scripts, for those
  nothing more needs be done. Others kernels are packages as notebooks, either
  as Rmarkdown or Jupyter, for those we use \c{knitr} to extract runnable code.
  The body of each extracted program is instrumented with calls to our dynamic
  analyzer to ensure that we only record calls to \eval from the code of
  interest and not from bootstrapping or execution harness operations.
\item \emph{Trace.} Each program is executing using our dynamic analysis tool
  which is a heavily instrumented interpreter that captures calls to \eval and
  many other fine-grained runtime events. Packages are run twice, once to
  capture \eval calls originating from package code, and a second time to
  capture calls coming from the base libraries. To avoid any interference, each
  program is run in its own process.
  As the GNU R bytecode compiler is written in R, we turn it off
  to avoid recording \eval in its code, and fallback on the interpreter.
\item \emph{Analyze.} Finally, analysis output is merged, cleaned and summarized
  in a post-processing phase driven by series of R scripts. The summarized data
  is then analyzed in RMarkdown notebooks to gather insights. All figures and
  numbers appearing in the paper are generated automatically. Figures are
  produced in PDF by \c{ggplot2}, numbers are exported as \LaTeX macros.
\end{compactenum}

@@

The tracer registers callbacks to all \eval functions as well as a few other
ones that allow us to better identify where the expressions come from. For
example, we taint the results of calls to \c{parse} (turns string into
expression) and \c{match.call} (reflects on current call). We also capture calls
to R API for dynamic code loading (\eg\xspace \c{library}, \c{require} and
\c{source}). Next, we subscribe to the events related to a variable definition
and assignment allowing us to record side effects that happen in environments
while evaluating code in \eval. One challenge we encountered during the analysis
was that essentially R only provides source references for top-level calls in a
function body or in the block surrounded by braces. That is to say, a method
whose body would \c{eval(x)} would not have debug information that we can use to
identify that particular \eval. The same expression surrounded by braces would
have a line number and source reference. This is unfortunately not easily fixed
in the R implementation. Instead, we extended the dynamic analysis tool to
attach synthetic source code references to all \eval call sites by traversing
ASTs. The tracer is implemented as an R package in 3.2K C++ and 1.3K of R code.
For performance reason, most of the tracing is done in C++. While in theory the
implementation is rather straightforward, it is not so in practice:
%
\begin{compactitem}[---]

\item The lazy evaluation makes it difficult to analyze function arguments
  while tracing as prematurely forcing a promise might have dangerous
  consequences.

@@

\item In the dynamic analysis, we run all the \emph{runnable} code we obtained
  from CRAN and Kaggle---\ie real code written by people with highly varying
  expertise in R and programming in general. This exercises a lot of the corner
  cases of the highly underspecified R behavior.

@@

\mypara{Limitations.} The analysis has some limitation. Even with the extension
described above, there are still \PkgUndefinedRnd \eval calls without source
references. This occurs when a reference to the \eval function is passed as
argument to a higher-order functions or when the \eval call originates from
native code. However, these missing source references account for a meager
\PkgUndefinedRatio of all calls; they are unlikely to affect our results. Other
limitations are that we ignore calls to the native \eval function and to the
alternate \c{rlang::tidy\_eval} function which uses native \eval internally.
The \c{rlang} package introduces a new kind of promise called \c{quosure} which
is evaluated by \c{tidy\_eval}. None of these limitations should invalidate our
conclusions.

>>>>

Our corpus is assembled from three sources: the \emph{Base} libraries 
bundled with the language implementation, packages hosted on \emph{CRAN}, and scripts from \emph{Kaggle}.

\mypara{Base.} The dataset contains \BasePackages libraries performing basic
arithmetics, statistics, and operating system functionalities. These libraries
are bundled with R and executed pervasively. Together they contain
\BaseEvalCallSites call sites to \eval in \BaseFunsWithEvals functions. Some of these
functions provide basic functionalities such as package loading, and thus, there
is hardly any code that does not invoke one of these.

\mypara{CRAN.} The Comprehensive R Archive Network
({\small \url{cran.r-project.org}}) is the largest curated repository of R
packages. It hosts \CranAvailablePackagesRnd packages with 6 new ones submitted
daily~\cite{Ligges2017}. Packages are authored by experienced developers and
abide by well-formedness rules automatically checked on each commit.
Each package comes with sample data. There are three sources of
runnable code in a package: \emph{tests, examples}, and \emph{vignettes} --
respectively, unit tests, code snippets from the documentation, and long-form
use-cases written in RMarkdown. Examples and vignettes can be turned into
scripts by extracting the relevant code. From \CranPackages packages we
extracted \CranRunnableScripts scripts that contain \CranRunnableCode lines of
code (\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd in
vignettes and \CranRunnableCodeTestsRnd in tests).

@@

The analysis presented in this has been automated by a pipeline that acquires
packages, extracts scripts, executes them, traces their behavior, and summarizes observations. Figure~\ref{fig:pipeline} shows the main steps along
 with their running time, data size, and  number of elements manipulated.
  Timings are from a cluster of three 2.3GHz Intel Xeon 6140 servers, with 72 cores and 256GB of RAM, and shared OCFS network storage. The pipeline steps are:

@@

\medskip
\begin{compactenum}
\item \emph{Download.} Packages are downloaded from CRAN. For scripts, a web
  crawler retrieves code, and the Kaggle command-line tool gets data.
  Installation is complicated by native dependencies, which are not adequately
  documented and thus hard to resolve automatically.
\item \emph{Extract.} Given installed packages and scripts, the next step is to
  create runnable programs. The \genthat tool helpfully extracts all runnable
  code snippets from a package and turns each of these into a self-standing
  program~\cite{issta18}. Some Kaggle kernels are already scripts. For those,
  nothing more needs to be done. Others kernels are packages as notebooks, either   as Rmarkdown or Jupyter. For those, we use \c{knitr} to extract
   runnable code.
  The body of each extracted program is instrumented with calls to our dynamic
  analyzer to ensure that we only record calls to \eval from the code of
  interest and not from bootstrapping or execution harness operations.
\item \emph{Trace.} Each program executes using our dynamic analysis tool, a heavily instrumented interpreter that captures calls to \eval and
  many other fine-grained runtime events. Packages are run twice, once to
  capture \eval calls originating from package code and a second time to
  capture calls coming from the base libraries. To avoid any interference, each
  program is run in its own process.
  As the GNU R bytecode compiler is written in R, we turn it off
  to avoid recording \eval in its code, and fallback on the interpreter.
\item \emph{Analyze.} Finally, the analysis output is merged, cleaned, and summarized
  in a post-processing phase driven by series of R scripts. The summarized data
  is then analyzed in RMarkdown notebooks to gather insights. All figures and
  numbers appearing in the paper are generated automatically. Figures are
  produced in PDF by \c{ggplot2}; numbers are exported as \LaTeX macros.
\end{compactenum}

@@

The tracer registers callbacks to all \eval functions as well as a few other
ones that allow us to identify better where the expressions come from. For
example, we taint the results of calls to \c{parse} (turns strings into
expressions) and \c{match.call} (reflects on current call). We also capture calls
to R API for dynamic code loading (\eg\xspace \c{library}, \c{require}, and
\c{source}). Next, we subscribe to the events related to a variable definition
and assignment, allowing us to record side effects that happen in environments
while evaluating code in \eval. One challenge we encountered during the analysis
was that essentially R only provides source references for top-level calls in a
function body or in the block surrounded by braces. That is to say, a method
whose body would \c{eval(x)} would not have debug information that we can use to
identify that particular \eval. The same expression surrounded by braces would
have a line number and source reference. This is unfortunately not easily fixed
in the R implementation. Instead, we extended the dynamic analysis tool to
attach synthetic source code references to all \eval call sites by traversing
ASTs. The tracer is implemented as an R package in 3.2K C++ and 1.3K of R code.
For performance reasons, most of the tracing is done in C++. While in theory, the
implementation is relatively straightforward, it is not so in practice:
%
\begin{compactitem}[---]

\item The lazy evaluation makes it challenging to analyze function arguments
  while tracing as prematurely forcing a promise might have dangerous
  consequences.

@@

\item In the dynamic analysis, we run all the \emph{runnable} code we obtained
  from CRAN and Kaggle---\ie actual code written by people with highly varying
  expertise in R and programming in general. This exercises a lot of the corner
  cases of the highly underspecified R behavior.

@@

\mypara{Limitations.} The analysis has some limitations. Even with the extension
described above, there are still \PkgUndefinedRnd \eval calls without source
references. This occurs when a reference to the \eval function is passed as an 
argument to a higher-order function or when the \eval call originates from
native code. However, these missing source references account for a meager
\PkgUndefinedRatio of all calls; they are unlikely to affect our results. Other
limitations are that we ignore calls to the native \eval function and to the
alternate \c{rlang::tidy\_eval} function, which uses native \eval internally.
The \c{rlang} package introduces a new kind of promise called \c{quosure} which
is evaluated by \c{tidy\_eval}. None of these limitations should invalidate our
conclusions.
